New orbits for irregular satellites based on thousands of new observations and predictions of stellar occultations up to 2020

### 2016-04-01 Irregular-occ/Irregular-occ.tex
.

\title[New orbits and predictions of stellar occultations by irregular satellites up to 2020]{New numerical integration of orbits of irregular satellites and prediction of stellar occultations up to 2020}
\author[A. R. Gomes-J\'unior, M. Assafin, L. Beauvalet et al.]{A. R. Gomes-J\'unior$^{1}$\thanks{E-mail: altair08@astro.ufrj.br},
M. Assafin$^{1,\dag}$,
L. Beauvalet$^{2,3}$,
J. Desmars$^{4}$,
R. Vieira-Martins$^{1,2,5,\dag}$,\newauthor
J. I. B. Camargo$^{2,5}$,
B. E. Morgado$^{1,2}$
F. Braga-Ribas$^{2,6}$
\\
$^{1}$Observat\'orio do Valongo/UFRJ, Ladeira Pedro Ant\^onio 43,
CEP 20.080-090 Rio de Janeiro - RJ, Brazil\\
$^{2}$Observat\'orio Nacional/MCTI, R. General Jos\'e Cristino 77, CEP 20921-400 Rio de Janeiro - RJ, Brazil\\
$^{3}$Observatoire de Paris/SYRTE, 77 Avenue Denfert Rochereau 75014 Paris, France\\
$^{4}$Institut de m\'ecanique c\'eleste et de calcul des \'eph\'em\'erides - Observatoire de Paris, UMR 8028 du CNRS, 77 Av. Denfert-Rochereau,\\ 75014 Paris, France\\
$^{5}$Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ 20921-400, Brazil\\
$^{6}$Federal University of Technology - Paran\'a (UTFPR / DAFIS), Rua Sete de Setembro, 3165, CEP 80230-901, Curitiba, PR, Brazil
}
\begin{document}
\newcommand{\noccs}{5457 } % total number of occultations

@@

\begin{abstract}
Due to their orbital configurations, it is common belief that the irregular satellites were probably captured by the giant planets in the early solar system. It is important to know their physical parameters, such as size, shape, albedo and composition, to trace back their true origin. The best ground-based technique to determine size and shape, and thus constrain the albedo and in a broader sense composition, is the observation of stellar occultations by these objects.
We aim to predict stellar occultations for the eight largest irregular satellites of Jupiter: Himalia, Elara, Pasiphae, Carme, Lysithea, Sinope, Ananke and Leda, and for the irregular satellites Phoebe of Saturn and Nereid of Neptune, and also for Triton.
We identified candidates to stellar occultations by the irregular satellites from the UCAC4 catalogue and from a catalogue of stars in the sky-path of Neptune obtained from observations made with the ESO2p2/WFI (2.2 m Max-Planck ESO telescope with the Wide Field Imager) instrument. These catalogues were crossed with the ephemeris of the satellites to identify stellar occultations. We used a new ephemeris based solely on the observations from \cite{GomesJunior2015} to generate predictions for the short-time future of the satellites of Jupiter. For Phoebe, we used an updated ephemeris and for Nereid and Triton we used recently published ephemeris.
We managed to identify \noccs candidates of stellar occultations between the period of January, 2016 and December, 2020. We made observational tests for the prediction of the event of Himalia in March 03, 2015. The stars and objects were observed close to the date predicted and in the same field to minimize errors and the obtained relative satellite-star positions were used to evaluate the predictions. The comparisons between the predictions and the observation tests show a good agreement. We call attention for an occultation by Triton of a bright star (R=12.5) in October 5, 2017. The event can be observed from Europe and the east coast of USA and will be a good opportunity to access the current state of the atmosphere of the satellite. We also note that Jupiter will cross the galactic plane in 2019-2020 and Saturn in 2018. This results in a large number of occultations predicted for that period.
We discuss how the successful observation of a stellar occultation by these objects is quite possible and present some of those potential occultations.
\end{abstract}

@@

Irregular satellites revolve around giant planets at large distances in eccentric, highly inclined and frequently retrograde orbits. Because of these peculiar orbits, it is largely accepted that these objects did not form by accretion around their planet, but were captured in the early solar system \citep{Sheppard2005}.

@@

In order to obtain precise fundamental physical parameters like size and shape, thus constraining the albedo and in a broader sense also the composition for the irregular satellites and therefore to contribute to the study of their origin, we aim at observing stellar occultations, which provide  more accurate results than other ground-based techniques \citep{Sicardy2011, Ortiz2012, Braga-Ribas2014}. For that, reliable predictions of stellar occultations by these satellites are most needed.

We present in this paper new numerical integration of the orbits of the 8 major irregular satellites of Jupiter (Himalia, Elara, Pasiphae, Lysithea, Carme, Ananke, Sinope and Leda) for predict stellar occultations. This is a specific short-time ephemeris using only the positions obtained by \cite{GomesJunior2015}. \cite{GomesJunior2015} obtained 6523 suitable positions for 18 irregular satellites between 1992 and 2014 with an estimated error in the positions of about 60 to 80 mas. For some satellites the number of positions obtained is comparable to the number used in the numerical integration of orbits by the JPL \citep{Jacobson2012}. They pointed out that the ephemeris of the irregular satellites have systematic errors that may reach 200 mas for some satellites. For an object at the distance of Jupiter, this represents an error langer than 700 km in the shadow path of an occultation.

For Phoebe, we updated the ephemeris of \cite{Desmars2013} using the observations of \cite{GomesJunior2015}, observations from Minor Planet Center and observations from Flagstaff. %For Nereid and Triton we used the most recent ephemeris available.

Also, with new and precise ephemeris, we present stellar occultation predictions for these 11 objects up to 2020. Phoebe, being the most studied object with a good measured size, can be used to calibrate and evaluate the technique for similar objects. Triton, although it is not an actual irregular satellite, was also include in the list because of its atmosphere which has a complicated and extreme seasonal cycle \citep{McKinnon2007, Elliot_2000}.

%Triton is an uncommon satellite. Its orbit is retrograde and inclined, but quasi-circular and very close to the planet compared to the irregular ones. Because Triton's orbit size is very small and its precession is not dominated by Solar perturbations, Triton is frequently excluded from the irregular satellites' class, but still studied together by many authors \citep{Sheppard2005, Jewitt2007}. Similarly to the irregular satellites, Triton was probably captured in the early solar system and may have the same origin as the TNOs \citep{Agnor2006}. However, Triton is bigger than the irregular satellites by an order of magnitude and has an atmosphere. The main motivation to study Triton by stellar occultations is to understand the evolution of its atmosphere due to Triton's complicated and extreme seasonal cycle \citep{McKinnon2007, Elliot_2000}. For all these reasons, Triton was also included as a target in this work.

Excluding Triton \citep{Olkin1997, Elliot_2000}, no observation of a stellar occultation by an irregular satellite was published up to date. Since their estimated sizes are very small (see Table \ref{Tab: satellite-diameter}), this may have discouraged earlier tries. But, in fact, given their relatively closer distances as compared to TNOs and Centaurs, and considering the current precision of their ephemeris and of star positions, we can now reliably predict the exact location and instant where the shadow of the occultation will cross the Earth. %For instance, Himalia, supposedly the largest irregular satellite of Jupiter has an estimated size of 150 km \citep{Porco2003}, which is equivalent to an apparent size of about 40 mas (milliarcseconds). Thus, in this case, if the accumulated error (2 sigma or 95\% confidence level) of ephemeris and star position be around 70 mas, we have a probability of about 30\% of observing a stellar occultation, which is quite satisfactory today (see discussion in section \ref{Sec: discussion}).

\begin{table}
\caption{\label{Tab: satellite-diameter} Estimated diameter of the satellites and correspondent apparent diameter}
\begin{center}
\begin{tabular}{lccc}
\hline  \hline
\multicolumn{4}{c}{Diameter of the satellites} \tabularnewline
Satellite  & mas$^ {a}$  & km & Ref. \tabularnewline
\hline
Ananke & 8 & 29 & 1 \tabularnewline
Carme & 13 & 46 & 1 \tabularnewline
Elara & 24 & 86 & 1 \tabularnewline
Himalia & 41 & $(150\times120) \pm 20^{b}$ & 2 \tabularnewline
Leda & 5 & 20 & 1 \tabularnewline
Lysithea & 10 & 36 & 1 \tabularnewline
Pasiphae & 17 & 62 & 1 \tabularnewline
Sinope & 10 & 37 & 1 \tabularnewline
\hdashline
Phoebe & 32 & $212 \pm 1.4^{b}$ & 3 \tabularnewline
\hdashline
Nereid & 15 & $340 \pm 50^{c}$ & 4 \tabularnewline
Triton & 124 & $2707 \pm 2.0^{c}$ & 5 \tabularnewline
\hline
\end{tabular}
\end{center}
References: (1) \cite{Rettig2001}; (2) \cite{Porco2003}; (3) \cite{Thomas2010}; (4) \cite{Thomas1991}; (5) \cite{Thomas2000}.\\
$^{a}${Using a mean distance from Jupiter of 5 AU, from Saturn of 9 AU and from Neptune of 30 AU.}
$^{b}${From Cassini observations.}
$^{c}${From Voyager-2 observations.}
\par
\end{table}

@@

In section \ref{Sec: integration} we show the building of the new ephemeris. In section \ref{Sec: predictions}, we present the predictions of the stellar occultations by irregular satellites and how they were made. Some tests made to check the accuracy of the predictions are presented in section \ref{Sec: testes} and the final discussion is presented in section \ref{Sec: discussion}.

\section{Ephemeris} \label{Sec: integration}

In order to improve ephemeris in short time spans in the near future for a more efficient prediction of stellar occultations by irregular satellites, we need predictions based on new and recent observations. In this context, \cite{GomesJunior2015} published 6523 precise positions for 18 irregular satellites from observations made at the Observatório do Pico dos Dias (OPD), Observatoire Haute-Provence (OHP) and European Southern Observatory (ESO) between 1992 and 2014.

Here we develop our own ephemeris based on the observations published in \cite{GomesJunior2015}. First, because the reduction was made with consistent and precise stellar catalogue and with a robust and reliable software \citep[PRAIA,][]{Assafin2011}. Second, besides recent observations, this consistent set of numerous and precise positions covers many orbital periods at many distinct orbital plane sights, so that the orbital inclinations along with all other orbital elements could be satisfactorily derived without the need of further position sets. For these reasons, only this set of positions was used for building the new ephemeris for the satellites of Jupiter.

@@

\subsection{Triton's and Nereid's ephemeris}

Triton is an uncommon satellite. Its orbit is retrograde and inclined, but quasi-circular and very close to the planet compared to the irregular ones. Because Triton's orbit size is very small and its precession is not dominated by Solar perturbations, Triton is frequently excluded from the irregular satellites' class, but still studied together by many authors \citep{Sheppard2005, Jewitt2007}. Similarly to the irregular satellites, Triton was probably captured in the early solar system and may have the same origin as the TNOs \citep{Agnor2006}. However, Triton is bigger than the irregular satellites by an order of magnitude and has an atmosphere.

For this satellite, we used the most recent ephemeris published by \cite{Emelyanov2015} which is based on more observations than previous ephemeris and covers 4 years more than the ephemeris of \cite{Jacobson2009}. In Fig. \ref{Fig:eph-Triton} we compare it with the ephemeris published in \cite{Zhang2014} and the NEP081 JPL ephemeris \citep{Jacobson2009}. The offsets between the three ephemeris are smaller than 20 mas for the period 2015-2021. This value is much smaller than the apparent size of Triton (124 mas, see Table \ref{Tab: satellite-diameter}), which indicates a good probability of observing an event by this object.

\begin{figure*}
\begin{centering}
\subfigure[NEP081 JPL and \protect\cite{Emelyanov2015}]{\includegraphics[width=8.7cm]{figures/JPL-EME_Triton.eps} \label{Fig:eph-Triton-eme}}
\subfigure[\protect\cite{Zhang2014} and \protect\cite{Emelyanov2015}]{\includegraphics[width=8.7cm]{figures/Zhang-EME_Triton.eps} \label{Fig:eph-Triton-zhang}}
\caption{Comparison between the \protect\cite{Emelyanov2015}, \protect\cite{Zhang2014} and NEP081 JPL \protect\citep{Jacobson2009} ephemeris for the satellite Triton.}
\label{Fig:eph-Triton}
\end{centering}
\end{figure*}

For Nereid, we used the ephemeris published by \cite{Emelyanov2011}, which uses more recent observations than the JPL ephemeris published by \cite{Jacobson2009}. The comparison between the two ephemeris can be seen in Fig. \ref{Fig:eph-Nereid}. The right ascension offsets between them are smaller than 60 mas and the declination ones are smaller than 15 mas. For Nereid, ephemeris errors in right ascension translate to uncertainties in the central instant of a stellar occultation, so offsets of 60 mas are not critical. The declination ephemeris offsets are of the order of the apparent diameter (15 mas, see Table \ref{Tab: satellite-diameter}), resulting in fair chances (50\%) that the shadow crosses the expected Earth latitude predicted for an occultation.

\begin{figure}
\begin{centering}
\includegraphics[width=8.8cm]{figures/JPL-EME_Nereid.eps}
\caption{Comparison between the \protect\cite{Emelyanov2011} and NEP081 JPL \protect\citep{Jacobson2009} ephemeris for the satellite Nereid.}
\label{Fig:eph-Nereid}
\end{centering}
\end{figure}

\cite{GomesJunior2015} published 902 new positions for Nereid from observations between 1992 and 2014. The development of a better ephemeris for Nereid using these new positions in the context discussed here of near future stellar occultations is desirable, but is out of the scope of this paper.

\section{Prediction of occultations} \label{Sec: predictions}

@@

We predicted occultations for the 8 major irregular satellites of Jupiter,  Ananke, Carme, Elara, Himalia, Leda, Lysithea, Pasiphae and Sinope, and for Phoebe of Saturn and Triton and Nereid of Neptune.

For Triton and Nereid, the candidates for stellar occultations in 2016 were searched using the WFI catalogue in the same way as the predictions for Centaurs and TNOs occultations by \cite{Assafin2010, Assafin2012} and \cite{Camargo2014}. This catalogue contains the stars in the path of Neptune in the sky up to mid-2016. The catalogue was generated by observations made at the ESO 2p2 telescope (IAU code 809) using the Wide Field Imager (WFI) CCD mosaic detector. The filter used was the broad-band R filter ESO\#844 with $\lambda_c$ = 651.725 nm and $\Delta\lambda$ = 162.184 nm.

A total of \noccs events were identified between January 2016 and December 2020. In Table \ref{Tab: satellite-occultation} we present the number of stellar occultations predicted by year for each satellite. It is possible to see an increase in the number of events found for Phoebe in 2018 and for the satellites of Jupiter in 2019-2020. This is because at that periods these satellites will cross the apparent galactic plane. In contrast, for Nereid and Triton, almost no occultation was found because Neptune is in a very low dense region of stars and move more slowly than the other planets. We call attention that about 10\% (516) of the events will involve stars brighter than magR=14 (and almost 25\% brighter than magR=15), which helps the attempt of amateur observers.

Table \ref{Tab: occ-list} shows a sample of the catalogue of occultations generated and their parameters, which are necessary to produce occultation maps. Since these objects are very small, the duration of each event is a few seconds. All the occultation tables and maps will be publicly available at the CDS. No star brighter than MagR*=18 will be occulted by Triton in 2016. For this satellite, we cut events with stars fainter than R=18, since for Triton the flux drop during such an occultation would be very small. In Fig. \ref{Fig: ocultacao} we show an example of an occultation map. This is one of the only two found occultation by Triton and it will happen in October 05, 2017. This event can be observed from Europe and the east coast of USA and will be a great opportunity to study in high resolution the atmosphere of the satellite.

\begin{table}
\caption{\label{Tab: satellite-occultation} Number of stellar occultations for each satellite from January, 2016 up to December, 2020.}
\begin{centering}
\begin{tabular}{lcccccc}
\hline  \hline
%\multicolumn{4}{c}{Diameter of the satellites} \tabularnewline
Satellite  & 2016 & 2017 & 2018 & 2019 & 2020 & Total \tabularnewline
\hline
Ananke & 12 & 16 & 49 & 359 & 187 & 623 \tabularnewline
Carme & 20 & 14 & 30 & 369 & 220 & 653 \tabularnewline
Elara & 14 & 16 & 33 & 305 & 193 & 561 \tabularnewline
Himalia & 15 & 12 & 54 & 257 & 230 & 568 \tabularnewline
Leda & 8 & 24 & 38 & 362 & 208 & 640 \tabularnewline
Lysithea & 16 & 11 & 35 & 330 & 212 & 604 \tabularnewline
Pasiphae & 20 & 19 & 44 & 362 & 206 & 651 \tabularnewline
Sinope & 15 & 21 & 34 & 356 & 256 & 682 \tabularnewline
\hdashline
Phoebe & 32 & 98 & 238 & 79 & 13 & 460 \tabularnewline
\hdashline
Nereid & 11$^{a}$ & 1 & 1 & 0 & 0 & 13 \tabularnewline
Triton & -- & 1 & 1 & 0 & 0 & 2 \tabularnewline
\hline
\end{tabular}
\par \end{centering}
$^{a}$ {Using the WFI catalogue as explained in Sec. \ref{Sec: predictions}}.
\end{table}

>>>>

\title[New orbits and predictions of stellar occultations by irregular satellites up to 2020]{New orbits for irregular satellites based on thousands of new observations and predictions of stellar occultations up to 2020}
\author[A. R. Gomes-J\'unior, M. Assafin, L. Beauvalet et al.]{A. R. Gomes-J\'unior$^{1}$\thanks{E-mail: altair08@astro.ufrj.br},
M. Assafin$^{1,\dag}$,
L. Beauvalet$^{2,3}$,
J. Desmars$^{4}$,
R. Vieira-Martins$^{1,2,5,\dag}$,\newauthor
J. I. B. Camargo$^{2,5}$,
B. E. Morgado$^{1,2}$
F. Braga-Ribas$^{2,6}$
\\
$^{1}$Observat\'orio do Valongo/UFRJ, Ladeira Pedro Ant\^onio 43,
CEP 20.080-090 Rio de Janeiro - RJ, Brazil\\
$^{2}$Observat\'orio Nacional/MCTI, R. General Jos\'e Cristino 77, CEP 20921-400 Rio de Janeiro - RJ, Brazil\\
$^{3}$Observatoire de Paris/SYRTE, 77 Avenue Denfert Rochereau 75014 Paris, France\\
$^{4}$Institut de m\'ecanique c\'eleste et de calcul des \'eph\'em\'erides - Observatoire de Paris, UMR 8028 du CNRS, 77 Av. Denfert-Rochereau,\\ 75014 Paris, France\\
$^{5}$Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ 20921-400, Brazil\\
$^{6}$Federal University of Technology - Paran\'a (UTFPR / DAFIS), Rua Sete de Setembro, 3165, CEP 80230-901, Curitiba, PR, Brazil
}
\begin{document}
\newcommand{\noccs}{5442 } % total number of occultations

@@

\begin{abstract}
%Due to their orbital configurations, it is common belief that the irregular satellites were probably captured by the giant planets in the early solar system. It is important to know their physical parameters, such as size, shape, albedo and composition, to trace back their true origin. The best ground-based technique to determine size and shape, and thus constrain the albedo and in a broader sense composition, is the observation of stellar occultations by these objects but this demands a very precise ephemeris based on new and good positions.
%\cite{GomesJunior2015} published 6523 observations obtained between 1992 and 2014 for 18 irregular satellites with an estimated error in the positions of about 60 to 80 mas.
%Based on this observations, we aimed to generate new numerical integration of the orbits of eight largest irregular satellites of Jupiter: Himalia, Elara, Pasiphae, Carme, Lysithea, Sinope, Ananke and Leda. This is a short-time ephemeris that corrected systematic errors from previous ephemeris (JPL JUP300). For Phoebe we updated ephemeris from \cite{Desmars2013} using 75\% more positions than the previous one.
%Also Jupiter will cross the galactic plane in 2019-2020 and Saturn in 2018. This creates a good opportunity to predict stellar occultations due to the large number of events that can be predicted in these periods. Thus, we aimed to predict stellar occultations for the same 9 satellites cited above. Using the UCAC4 catalogue we managed to identify \noccs candidates of stellar occultations between the period of January, 2016 and December, 2020.
%We discuss how the successful observation of a stellar occultation by these objects is quite possible and present some of those potential occultations.
\cite{GomesJunior2015} published 6523 new positions for 18 irregular satellites from observations made between 1992 and 2014 with an estimated error of about 60 to 80 mas. Based on this set of original positons, we performed numerical integrations for deriving new orbits for the eight largest irregular satellites of Jupiter: Himalia, Elara, Pasiphae, Carme, Lysithea, Sinope, Ananke and Leda. This corrected systematic errors and resulted in an ephemeris more accurate/precise than JPL JUP300 for these bodies at least in the immediate future. For Phoebe we updated the ephemeris from \cite{Desmars2013} using 75\% more positions than the previous one. Due to their orbital configurations, it is common belief that the irregular satellites were captured by the giant planets in the early Solar System, but there is no consensus for a single model explaining where they were formed. Size, shape, albedo and composition would help a lot to trace back their true origin, but these physical parameters are yet poorly known for irregular satellites. The observation of stellar occultations would allow for the determination of such parameters. Indeed Jupiter will cross the galactic plane in 2019-2020 and Saturn in 2018, improving a lot the chances of observing such events in the near future. Using the derived ephemerides and the UCAC4 catalogue we managed to identify \noccs candidate stellar occultations between January 2016 and December 2020 for the 9 irregular satellites studied here. We discussed how the successful observation of a stellar occultation by these objects is quite possible and present some of those potential occultations.
\end{abstract}

@@

Irregular satellites revolve around giant planets at large distances in eccentric, highly inclined and frequently retrograde orbits. Because of these peculiar orbits, it is largely accepted that these objects did not form by accretion around their planet, but were captured in the early Solar System \citep{Sheppard2005}.

@@

\cite{GomesJunior2015} obtained 6523 suitable positions for 18 irregular satellites between 1992 and 2014 with an estimated error in the positions of about 60 to 80 mas. For some satellites the number of positions obtained is comparable to the number used in the numerical integration of orbits by the JPL \citep{Jacobson2012}. They pointed out that the ephemeris of the irregular satellites have systematic errors that may reach 200 mas for some satellites.

We present in this paper new numerical integration of the orbits of the 8 major irregular satellites of Jupiter (Himalia, Elara, Pasiphae, Lysithea, Carme, Ananke, Sinope and Leda) using only the positions obtained by \cite{GomesJunior2015}. We aim to eliminate the systematic errors in the JPL ephemeris (see Sec. \ref{Sec: integration}). %For an object at the distance of Jupiter, this represents an error langer than 700 km in the shadow path of an occultation.
For Phoebe, we updated the ephemeris of \cite{Desmars2013} using the observations of \cite{GomesJunior2015}, \cite{Peng2015}, observations from Minor Planet Center and observations from Flagstaff.

Phoebe, being the most studied object with a good measured size, can be used to calibrate and evaluate the technique for similar objects. Up to date, no observation of a stellar occultation by an irregular satellite was published. Since their estimated sizes are very small (see Table \ref{Tab: satellite-diameter}), this may have discouraged earlier tries. But, in fact, given their relatively closer distances as compared to TNOs and Centaurs, and considering the precision of our ephemeris and of star positions, we can now reliably predict the exact location and instant where the shadow of the occultation will cross the Earth.

\begin{table}
\caption{\label{Tab: satellite-diameter} Estimated diameter of the satellites and correspondent apparent diameter}
\begin{center}
\begin{tabular}{lccc}
\hline  \hline
\multicolumn{4}{c}{Diameter of the satellites} \tabularnewline
Satellite  & mas$^ {a}$  & km & Ref. \tabularnewline
\hline
Ananke & 8 & 29 & 1 \tabularnewline
Carme & 13 & 46 & 1 \tabularnewline
Elara & 24 & 86 & 1 \tabularnewline
Himalia & 41 & $(150\times120) \pm 20^{b}$ & 2 \tabularnewline
Leda & 5 & 20 & 1 \tabularnewline
Lysithea & 10 & 36 & 1 \tabularnewline
Pasiphae & 17 & 62 & 1 \tabularnewline
Sinope & 10 & 37 & 1 \tabularnewline
\hdashline
Phoebe & 32 & $212 \pm 1.4^{b}$ & 3 \tabularnewline
%\hdashline
%Nereid & 15 & $340 \pm 50^{c}$ & 4 \tabularnewline
%Triton & 124 & $2707 \pm 2.0^{c}$ & 5 \tabularnewline
\hline
\end{tabular}
\end{center}
References: (1) \cite{Rettig2001}; (2) \cite{Porco2003}; (3) \cite{Thomas2010}; (4) \cite{Thomas1991}; (5) \cite{Thomas2000}.\\
$^{a}${Using a mean distance from Jupiter of 5 AU, from Saturn of 9 AU and from Neptune of 30 AU.}
$^{b}${From Cassini observations.}
$^{c}${From Voyager-2 observations.}
\par
\end{table}

@@

In section \ref{Sec: integration} the new determination of the orbits. In section \ref{Sec: predictions}, we present the predictions of the stellar occultations by irregular satellites, including some tests made to check the accuracy of the predictions. The final discussion is presented in section \ref{Sec: discussion}.

\section{Orbit computations} \label{Sec: integration}

\cite{GomesJunior2015} published 6523 precise positions for 18 irregular satellites from observations made at the Observatório do Pico dos Dias (OPD), Observatoire Haute-Provence (OHP) and European Southern Observatory (ESO) between 1992 and 2014.

Here we compute new orbits based on the observations published in \cite{GomesJunior2015}. First, because the reduction was made with a consistent and precise stellar catalogue and with a robust astrometry \citep[PRAIA,][]{Assafin2011}. Second, besides recent observations, this consistent set of numerous and precise positions covers many orbital periods at many distinct orbital plane sights, so that the orbital inclinations along with all other orbital elements could be satisfactorily derived without the need of further position sets. For these reasons, only this set of positions was used for the satellites of Jupiter.

Due to the context of this work regarding to stellar occultations, the orbit fitting procedures used aimed primarily to derive precise ephemerides for the near future. The procedures allow for the improvement of orbits as new observations are added.

@@

%\subsection{Triton's and Nereid's ephemeris}
%
%Triton is an uncommon satellite. Its orbit is retrograde and inclined, but quasi-circular and very close to the planet compared to the irregular ones. Because Triton's orbit size is very small and its precession is not dominated by Solar perturbations, Triton is frequently excluded from the irregular satellites' class, but still studied together by many authors \citep{Sheppard2005, Jewitt2007}. Similarly to the irregular satellites, Triton was probably captured in the early solar system and may have the same origin as the TNOs \citep{Agnor2006}. However, Triton is bigger than the irregular satellites by an order of magnitude and has an atmosphere.
%
%For this satellite, we used the most recent ephemeris published by \cite{Emelyanov2015} which is based on more observations than previous ephemeris and covers 4 years more than the ephemeris of \cite{Jacobson2009}. In Fig. \ref{Fig:eph-Triton} we compare it with the ephemeris published in \cite{Zhang2014} and the NEP081 JPL ephemeris \citep{Jacobson2009}. The offsets between the three ephemeris are smaller than 20 mas for the period 2015-2021. This value is much smaller than the apparent size of Triton (124 mas, see Table \ref{Tab: satellite-diameter}), which indicates a good probability of observing an event by this object.
%
%\begin{figure*}
%\begin{centering}
%\subfigure[NEP081 JPL and \protect\cite{Emelyanov2015}]{\includegraphics[width=8.7cm]{figures/JPL-EME_Triton.eps} \label{Fig:eph-Triton-eme}}
%\subfigure[\protect\cite{Zhang2014} and \protect\cite{Emelyanov2015}]{\includegraphics[width=8.7cm]{figures/Zhang-EME_Triton.eps} \label{Fig:eph-Triton-zhang}}
%\caption{Comparison between the \protect\cite{Emelyanov2015}, \protect\cite{Zhang2014} and NEP081 JPL \protect\citep{Jacobson2009} ephemeris for the satellite Triton.}
%\label{Fig:eph-Triton}
%\end{centering}
%\end{figure*}
%
%For Nereid, we used the ephemeris published by \cite{Emelyanov2011}, which uses more recent observations than the JPL ephemeris published by \cite{Jacobson2009}. The comparison between the two ephemeris can be seen in Fig. \ref{Fig:eph-Nereid}. The right ascension offsets between them are smaller than 60 mas and the declination ones are smaller than 15 mas. For Nereid, ephemeris errors in right ascension translate to uncertainties in the central instant of a stellar occultation, so offsets of 60 mas are not critical. The declination ephemeris offsets are of the order of the apparent diameter (15 mas, see Table \ref{Tab: satellite-diameter}), resulting in fair chances (50\%) that the shadow crosses the expected Earth latitude predicted for an occultation. 
%
%\begin{figure}
%\begin{centering}
%\includegraphics[width=8.8cm]{figures/JPL-EME_Nereid.eps}
%\caption{Comparison between the \protect\cite{Emelyanov2011} and NEP081 JPL \protect\citep{Jacobson2009} ephemeris for the satellite Nereid.}
%\label{Fig:eph-Nereid}
%\end{centering}
%\end{figure}
%
%\cite{GomesJunior2015} published 902 new positions for Nereid from observations between 1992 and 2014. The development of a better ephemeris for Nereid using these new positions in the context discussed here of near future stellar occultations is desirable, but is out of the scope of this paper.

\section{Prediction of stellar occultations} \label{Sec: predictions}

\subsection{Candidate events}

@@

We predicted occultations for the 8 major irregular satellites of Jupiter,  Ananke, Carme, Elara, Himalia, Leda, Lysithea, Pasiphae and Sinope, and for Phoebe of Saturn.

A total of \noccs events were identified between January 2016 and December 2020. In Table \ref{Tab: satellite-occultation} we present the number of stellar occultations predicted by year for each satellite. It is possible to see an increase in the number of events found for Phoebe in 2018 and for the satellites of Jupiter in 2019-2020. This is because at that periods these satellites will cross the apparent galactic plane. We call attention that about 10\% of the events will involve stars brighter than magR=14 (and almost 25\% brighter than magR=15), which helps the attempt of amateur observers.

\begin{table}
\caption{\label{Tab: satellite-occultation} Number of stellar occultations for each satellite from January, 2016 up to December, 2020.}
\begin{centering}
\begin{tabular}{lcccccc}
\hline  \hline
%\multicolumn{4}{c}{Diameter of the satellites} \tabularnewline
Satellite  & 2016 & 2017 & 2018 & 2019 & 2020 & Total \tabularnewline
\hline
Ananke & 12 & 16 & 49 & 359 & 187 & 623 \tabularnewline
Carme & 20 & 14 & 30 & 369 & 220 & 653 \tabularnewline
Elara & 14 & 16 & 33 & 305 & 193 & 561 \tabularnewline
Himalia & 15 & 12 & 54 & 257 & 230 & 568 \tabularnewline
Leda & 8 & 24 & 38 & 362 & 208 & 640 \tabularnewline
Lysithea & 16 & 11 & 35 & 330 & 212 & 604 \tabularnewline
Pasiphae & 20 & 19 & 44 & 362 & 206 & 651 \tabularnewline
Sinope & 15 & 21 & 34 & 356 & 256 & 682 \tabularnewline
\hdashline
Phoebe & 32 & 98 & 238 & 79 & 13 & 460 \tabularnewline
%\hdashline
%Nereid & 11$^{a}$ & 1 & 1 & 0 & 0 & 13 \tabularnewline
%Triton & -- & 1 & 1 & 0 & 0 & 2 \tabularnewline
\hline
\end{tabular}
\par \end{centering}
$^{a}$ {Using the WFI catalogue as explained in Sec. \ref{Sec: predictions}}.
\end{table}

Table \ref{Tab: sample-cds} shows a sample of the catalogue of occultations generated and their parameters, which are necessary to produce occultation maps. Since these objects are very small, the duration of each event is a few seconds. All the occultation tables and maps will be publicly available at the CDS. In Fig. \ref{Fig: ocultacao} we show an example of an occultation map. This is an occultation by Elara that will happen in February 21, 2017. This event can be observed from Australia and it is one of the best opportunities for this object due to the slow velocity of the event and it involves a bright star (MagR*=12.4).

\begin{table*}
\caption{\label{Tab: sample-cds} A sample of stellar occultation predictions for Pasiphae}
\begin{center}
\begin{tabular}{cccrcccrcrr}
\hline
\hline
d m Year~~~~h~~m~~~s & RA~~~(ICRS)~~~Dec & C/A & P/A & $\nu$ & {\it D} & $R^*$ & $\lambda$ & LST & $\mu_{\alpha*}$ & $\mu_{\delta}$ \tabularnewline
\hline
% & \multicolumn{2}{c}{Mean errors} & Nr & Nr &  UCAC4       \tabularnewline
09 04 2016 03:58:19. & 11 14 36.7707 +07 39 20.7610 & 1.003 &  17.9 & -12.88 &  4.54 & 14.9 & 271. & 22:03 &  12. & -33. \tabularnewline 
13 06 2016 00:16:12. & 11 12 48.5020 +07 06 43.3520 & 0.661 &  30.0 & +14.32 &  5.50 & 13.9 & 262. & 17:45 &  -1. &   1. \tabularnewline 
27 06 2016 13:56:09. & 11 18 03.4160 +06 23 45.1940 & 1.707 &  28.0 & +20.29 &  5.74 & 11.7 &  44. & 16:53 &   4. & -10. \tabularnewline 
18 07 2016 15:07:24. & 11 28 15.5076 +05 05 31.8060 & 0.942 &  26.7 & +27.80 &  6.05 & 14.0 &   8. & 15:40 &   4. &   4. \tabularnewline 
22 07 2016 16:15:07. & 11 30 30.4310 +04 48 43.4340 & 0.644 & 206.5 & +29.04 &  6.11 & 14.6 & 348. & 15:27 &  23. & -24. \tabularnewline 
24 07 2016 01:37:34. & 11 31 17.8471 +04 42 49.0540 & 0.029 & 206.6 & +29.46 &  6.12 & 15.1 & 206. & 15:22 &   2. &  -8. \tabularnewline 
24 07 2016 17:37:18. & 11 31 40.7472 +04 39 57.5060 & 0.840 &  26.5 & +29.66 &  6.13 & 14.9 & 326. & 15:20 & -11. &  -1. \tabularnewline 
\hline
\end{tabular}
\end{center}
\begin{flushleft}
\textbf{Notes}. Entries included: day of the year and UTC time of the prediction; right ascension and declination of the occulted star - at the central instant of occultation (corrected by proper motions); C/A: the geocentric closest approach, in arcseconds; P/A: the satellite position angle with respect to the occulted star at C/A, in degrees (zero at north of the star, increasing clockwise); $\nu$: velocity in the plane of sky, in km s$^{-1}$: positive = prograde, negative = retrograde; {\it D}: planet range to Earth, in AU; $R^*$: normalized magnitude to a common shadow velocity of 20 km s$^{-1}$ by the relationship $\textrm{R}^* = \textrm{R}_{\textrm{actual}} + 2.5 \times \log 10 \left(\frac{\textrm{velocity}} {20 \textrm{km}\, \textrm{s}^{-1}} \right)$; $\lambda$: east longitude of subplanet point in degrees, positive towards east; LST: UT + $\lambda$: local solar time at subplanet point, hh:mm; $\mu_{\alpha *}$ and $\mu_{\delta}$: proper motions in right ascension and declination, respectively (mas/year).
\end{flushleft}
\end{table*}
